$\pi$SOD-M: A Methodology for Building Service-Oriented Applications in the Presence of Non-Functional Properties

### 2014-03-15 JournalToSubmit2013/modelingWithPISODM.tex
New version with only necesary figures ... still undergoing

Given  (composite) $\pi$-use case model defining the functions of  the developed application and their associated constraints (representing business rules), the objective is to express them as a $\pi$-service process model describing the application logic and its associated non-functional properties.  The principle is to express a set of $\pi$-use cases (i.e., use cases with constraints) related by relationships of type {\sc Extend}   in terms of  a business process consisting of actions related by a control flow and associated contracts specifying NFP. The PIM to PIM transformation rules specify how  to transform a $\pi$-use case model into  a $\pi$-service process model (Table \ref{fig:transformationsUseCase-ServiceProcess}).

\begin{figure}
\centering{
\includegraphics[width=0.96\textwidth]{figs/Table5}}
\caption{ $\pi$-UseCase to $\pi$-ServiceProcess transformation}
\label{fig:transformationsUseCase-ServiceProcess}
\end{figure}

@@

\begin{example}[To Publish Music \textit{(cont)}]\label{ex:toPublicMusicT1}
Considering the scenario example  the transformation of the "listen music" Use Case is transformed into a Service Action. This Action is a Spotify service function that can be invoked to play the music. For the "publish music" Use Case,  constraints are transformed in a set of assertions that are grouped in a Contract ("publishMusicContract"). The "publishMusicContract" is related with the Action "publishMusic". 
For the use case "download music" it is necessary process  the payment. Thus, the use cases that model the function "download music"  are transformed into {\sf actions}, and a {\sf Service activity} that aggregates these {\sf Actions}. 
Considering the scenario example, it is possible to apply the same rule for the "publish music" use case, which has two extended use cases, "public twitter" and "public Facebook", applying the rules shown in figures \ref{fig:transformaton-examples}-b and d.
Given the "download music" use case of the scenario example, it includes the payment process to buy the music. It is represented as a include in the $\pi$-UseCase model and it is transformed in a sequence flow in the $\pi$-Service process model (as shown in figure \ref{fig:transformation-example-include}). 
\end{example}

\begin{figure}
\centering{
\includegraphics[width=0.76\textwidth]{figs/38}}
\caption{ Include Transformation Example}
\label{fig:transformation-example-include}
\end{figure}

@@

%Figure \ref{fig:pewsexpression} shows the correspondence between the model and the statements that implement it, with a schematic representation of the business process.
%\begin{figure}
%\centering{
%\includegraphics[width=0.85\textwidth]{figs/pews-expression}}
%\caption{Pews program implementing the "To Publish Music" application}
%\label{fig:pewsexpression}
%\end{figure}
Table \ref{fig:ServiceProcess-ServiceComposition}  shows the transformation rules for transforming a $\pi$-Service process model into a   $\pi$-ServiceComposition model.
As illustrated in Figure \ref{fig:ServiceProcess-ServiceComposition-Rules} the  principle of this transformation  is to group respectively  {\sf Contracts}  and {\sf Actions} into {\sf Policies} and {\sf Service activities}.

@@

\begin{figure}
\centering{
\includegraphics[width=1\textwidth]{figs/ServiceProcess-ServiceComposition}}
\caption{ Transformation Rules: From $\pi$-ServiceProcess to $\pi$-ServiceComposition}
\label{fig:ServiceProcess-ServiceComposition}
\end{figure}

@@

Table \ref{fig:ServiceComposition-Pews-Rules} describes the PIM to PSM transformations between a $\pi$-Service composition and a $\pi$-PEWS program. We propose two groups of rules: those that transform services composition entities of a $\pi$-Service composition model into $\pi$-PEWS model ; and those that transform rules grouped by policies into A-policy types.

@@

\begin{figure}
\centering{
\includegraphics[width=0.96\textwidth]{figs/Table7}}
\caption{Transformation Rules: From $\pi$-ServiceComposition to $\pi$-PEWS}
\label{fig:ServiceComposition-Pews-Rules}
\end{figure}

>>>>

Given  (composite) $\pi$-use case model defining the functions of  the developed application and their associated constraints (representing business rules), the objective is to express them as a $\pi$-service process model describing the application logic and its associated non-functional properties.  The principle is to express a set of $\pi$-use cases (i.e., use cases with constraints) related by relationships of type {\sc Extend}   in terms of  a business process consisting of actions related by a control flow and associated contracts specifying NFP. The PIM to PIM transformation rules specify how  to transform a $\pi$-use case model into  a $\pi$-service process model.
%(Table \ref{fig:transformationsUseCase-ServiceProcess}).

%\begin{figure}
%\centering{
%\includegraphics[width=0.96\textwidth]{figs/Table5}}
%\caption{ $\pi$-UseCase to $\pi$-ServiceProcess transformation}
%\label{fig:transformationsUseCase-ServiceProcess}
%\end{figure}

@@

\begin{example}[To Publish Music \textit{(cont)}]\label{ex:toPublicMusicT1}
Considering the scenario example  the transformation of the "listen music" Use Case is transformed into a Service Action. This Action  represents a Spotify service function that can be invoked to play the music. For the "publish music" use case,  constraints are transformed in a set of assertions that are grouped in a Contract ({\sf "publishMusicContract"}) associated to the Action {\sf "publishMusic"}. 
The "download music" use case  includes the payment process to buy the music. Thus, these use cases  are transformed into {\sf Actions}, and a {\sf Service activity} that aggregates these {\sf Actions}.   It is transformed in a sequence flow in the $\pi$-service process model (as shown in figure \ref{fig:transformation-example-include}). 
The same rule is applied for the "publish music" use case, which has two extended use cases, "public twitter" and "public Facebook" (figures \ref{fig:transformaton-examples}-b and d).
 \end{example}

%\begin{figure}
%\centering{
%\includegraphics[width=0.76\textwidth]{figs/38}}
%\caption{ Include Transformation Example}
%\label{fig:transformation-example-include}
%\end{figure}

@@

%Figure \ref{fig:pewsexpression} shows the correspondence between the model and the statements that implement it, with a schematic representation of the business process.
%\begin{figure}
%\centering{
%\includegraphics[width=0.85\textwidth]{figs/pews-expression}}
%\caption{Pews program implementing the "To Publish Music" application}
%\label{fig:pewsexpression}
%\end{figure}
%Table \ref{fig:ServiceProcess-ServiceComposition}  shows the transformation rules for transforming a .
As illustrated in Figure \ref{fig:ServiceProcess-ServiceComposition-Rules} the  principle of the transformation  of a  $\pi$-Service process model into a   $\pi$-ServiceComposition model is to group respectively  {\sf Contracts}  and {\sf Actions} into {\sf Policies} and {\sf Service activities}.

@@

%\begin{figure}
%\centering{
%\includegraphics[width=1\textwidth]{figs/ServiceProcess-ServiceComposition}}
%\caption{ Transformation Rules: From $\pi$-ServiceProcess to $\pi$-ServiceComposition}
%\label{fig:ServiceProcess-ServiceComposition}
%\end{figure}

@@

%Table \ref{fig:ServiceComposition-Pews-Rules} 
This section describes the PIM to PSM transformations between a $\pi$-Service composition and a $\pi$-PEWS program. We propose two groups of rules: those that transform services composition entities of a $\pi$-Service composition model into $\pi$-PEWS model ; and those that transform rules grouped by policies into A-policy types.

@@

%\begin{figure}
%\centering{
%\includegraphics[width=0.96\textwidth]{figs/Table7}}
%\caption{Transformation Rules: From $\pi$-ServiceComposition to $\pi$-PEWS}
%\label{fig:ServiceComposition-Pews-Rules}
%\end{figure}
